---
title: "Lab 1.1"
author: "Ron Guymon"
date: "2023-04-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
library(tidyverse) # For data wrangling
library(magrittr) # Pipe function
library(lubridate) # For working with dates

```

```{r}
monthlyProfit <- function(lawnsPerMonth = 200, chiefHours = 1){
  pricePerLawn <- 50
  truckPaymentPerMonth <- 300
  mowers <- 3
  mowerPaymentPerMonth <- 100
  equipmentCost <- 2000
  equipmentLifeInMonths <- 20
  equipmentMaintPerMonth <- 100
  insurancePerMonth <- 300
  fuelGallonsPerLawn <- 1
  fuelCostPerGallon <- 3
  suppliesPerLawn <- 1.5
  advertPerMonth <- ifelse(lawnsPerMonth ==200,50,300)
  chiefHourlyRate <- 15
  staffHourlyRate <- 10
  
  laborExpensePerLawn <- (chiefHourlyRate *chiefHours) + (staffHourlyRate*1)
  revenuePerMonth <- pricePerLawn*lawnsPerMonth
  expensePerMonth <- sum(truckPaymentPerMonth
                         , mowers*mowerPaymentPerMonth
                         , equipmentCost/equipmentLifeInMonths
                         , equipmentMaintPerMonth
                         , insurancePerMonth
                         , fuelGallonsPerLawn*fuelCostPerGallon*lawnsPerMonth
                         , suppliesPerLawn*lawnsPerMonth
                         , advertPerMonth
                         , laborExpensePerLawn*lawnsPerMonth
  )
  incomePerMonth <- revenuePerMonth - expensePerMonth
  incomePerLawn <- incomePerMonth / lawnsPerMonth
  
  returnList <- list(incomePerMonth = incomePerMonth
                     , laborExpensePerLawn = laborExpensePerLawn
                     , incomePerLawn = incomePerLawn)
  return(returnList)
  
}

```

# Scenarios 1 & 2: 200 lawns per month

```{r}
allProfit <- data.frame()
for(ch in seq(0, 1, .1)){
  mp <- monthlyProfit(lawnsPerMonth = 200, chiefHours = ch)
  tdf <- data.frame(laborExpensePerLawn = mp$laborExpensePerLawn
                    , monthlyProfit = mp$incomePerMonth)
  
  allProfit %<>% bind_rows(tdf)
}

# Plot
ggplot(allProfit, aes(x = laborExpensePerLawn, y = monthlyProfit)) +
  geom_point() +
  theme_minimal() +
  labs(title = 'Monthly Profit by Labor Cost Per Lawn'
       , x = 'Labor Cost Per Lawn'
       , y = 'Monthly Profit')

```

__Conclusion__: Yes, as long as the chief mower does not take more than 1 hour per lawn, the business will be profitable. The monthly profit ranges from about \$2,900 when the labor per lawn is \$25 to about \$6,000 when the labor per lawn is \$10.


# Scenarios 3 & 4: 350 lawns per month

```{r}
allProfit$scenario <- '1 & 2 (200 lawns)'
for(ch in seq(0, 1, .1)){
  mp <- monthlyProfit(lawnsPerMonth = 350, chiefHours = ch)
  tdf <- data.frame(laborExpensePerLawn = mp$laborExpensePerLawn
                    , monthlyProfit = mp$incomePerMonth
                    , scenario = '3 & 4 (350 lawns)')
  
  allProfit %<>% bind_rows(tdf)
}

# Plot
ggplot(allProfit, aes(x = laborExpensePerLawn, y = monthlyProfit)) +
  geom_point(aes(color = scenario)) +
  theme_minimal() +
  labs(title = 'Monthly Profit by Labor Cost Per Lawn'
       , x = 'Labor Cost Per Lawn'
       , y = 'Monthly Profit')

```

__Conclusion__: Yes, as long as the chief mower does not take more than 1 hour per lawn, the business will be profitable. The monthly profit ranges from about \$5,800 when the labor per lawn is \$25 to about \$11,000 when the labor per lawn is \$10.  

The company is much more profitable when they mow 350 lawns.
